Highly-Available Functional Replicated Datatypes

### 2017-02-21 paper/model.tex
Model edits.

\label{fig:canvas-sig}
\caption{\drawsome: a sample \name application}
\end{figure}

@@

Fig.~\ref{canvas-sig} shows the signature of the \drawsome
application. \drawsome represents a free-hand drawing canvas as a tree
of quadrants, where each quadrant is simply a leaf node unless it has
pixels of different colors. Quadrants are expanded into a tree
structures as an when pixels are colored (the color scheme used is
rgb). The representation is thus optimized for sparse canvases, such
as whiteboards. The application supports three simple operations:
creating a new canvas, setting the pixel at a specified coordinate,
and returning the pixel at a given coordinate. The \C{merge} function
and \C{deriving versioned} directive are explained below.

@@

The \drawsome example demonstrates the utility of mergeable data types
and the \name programming model in building concurrent applications
with conceptual sharing of state. This example, however, implicitly
makes certain assumptions about the model and the underlying system,
such as the existence of a single least common ancestor (LCA) for any
pair of versions, the ability to access a previous version on any
branch, and mergeability of any two concurrent versions. Enforcing
these guarantees in a fully decentralized distributed setting requires
addressing non-trivial theoretical challenges described in the next
section. Unlike the canvas, the process of writing merge functions for
data types with non-trivial invariants could be quite involved. The
later sections describe a principled approach to derive a merge
function for a data type given its invariants. Like \C{Canvas.merge},
merge functions often need to compare multiple versions of a data type
for structural equality, which, when performed naively, incurs cost
linear in the size of versions. The representation of a versioned data
type in \name is optimized for performing comparisons such that
structural equality can be determined in constant time. Such
engineering details, which contribute to the practicality of our
programming model are also discussed in subsequent sections.

>>>>

\caption{\drawsome: a sample \name application}
\label{fig:canvas-sig}
\end{figure}

@@

Fig.~\ref{fig:canvas-sig} shows the signature of the \drawsome
application. \drawsome represents a free-hand drawing canvas as a tree
of quadrants, where each quadrant is simply a leaf node unless it has
pixels of different colors. Quadrants are expanded into a tree
structures as an when pixels are colored (the color scheme used is
rgb). The representation is thus optimized for sparse canvases, such
as whiteboards. The application supports three simple operations:
creating a new canvas, setting the pixel at a specified coordinate,
and returning the pixel at a given coordinate. The \C{merge} function
and \C{deriving versioned} directive are explained below.

@@

The \drawsome example demonstrates the utility of mergeable data types
and the \name programming model in building concurrent applications
with conceptual sharing of state. This example, however, implicitly
makes certain assumptions about the model and the underlying system,
such as the existence of a single least common ancestor (LCA) for any
pair of versions, the ability to access a previous version on any
branch, and mergeability of any two concurrent versions. Enforcing
these guarantees in a fully decentralized distributed setting requires
addressing non-trivial theoretical challenges described in the next
section. Unlike the canvas, the process of writing merge functions for
data types with non-trivial invariants could be quite involved. The
later sections describe a principled approach to derive a merge
function for a data type given its invariants. Like \C{Canvas.merge},
merge functions often need to compare multiple versions of a data type
for structural equality, which, when performed naively, incurs cost
linear in the size of versions. The representation of a versioned data
type in \name is optimized for performing comparisons that allows
structural equality to be determined in constant time. Such
engineering details, which contribute to the practicality of our
programming model are also discussed in subsequent sections.
